25 Years of Water Quality Change in Rhode Island Lakes and Ponds

### 2019-01-07 manuscript/manuscript.Rmd
Add paragraph describing URI WW
Focuses on lakes & ponds

>>>>

The University of Rhode Island's Watershed Watch (URI WW) Volunteer Monitoring program has been collecting water quality data from Rhode Island lakes and ponds for over 25 years. The program began in 1988, monitoring 14 lakes and has now grown to include over 250 monitoring sites on over 120 waterbodies, including rivers/streams, and estuaries, with more than 400 trained volunteers. URI WW now provides more than 90% of Rhode Island's lake multi-year baseline data, and is an integral part of the state's environmental data collection strategy. Data QA/QC is of paramount importance; volunteers are trained both in the classroom and the field, and are provided with all the necessary equipment and supplies, along with scheduled collection dates. For freshwater lakes and ponds, weekly secchi depth and water temperature are recorded, along with bi-weekly Chl-a and dissolved oxygen. Water samples are collected three times per season (May through October) to be analyzed in the EPA-certified laboratory for nutrients, alkalinity, pH, and bacteria.
